---
title: The dynamic landscape of peptide function prediction
output:                                                                                                         
  html_document:
    theme: paper                                                                               
---      

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(DT)
library(dplyr)

dat <- readRDS("../data/curated_dat.RDS") %>% 
  filter(isPEP == TRUE & isML == TRUE) %>% 
  mutate(DOI = case_when(!is.na(publication) ~ publication,
                         !is.na(preprint) ~ preprint,
                         !is.na(thesis) ~ thesis),
         Type = case_when(!is.na(publication) ~ "Publication",
                         !is.na(preprint) ~ "Pre-print",
                         !is.na(thesis) ~ "Thesis")) %>% 
  select(c("name", "DOI", "Type", "reproducibility", "webserver", "webserveractive", "modelrepository", "trainingrepository",
           "antiangiogenic", "antibacterial", "antibiofilm", "anticancer", "antifungal", "antihypertensive", "antiinflammatory",
           "antimicrobial", "antiparasitic", "antiviral", "blood-brain barrier", "cell-penetrating", "chemotactic", "neuropeptides",
           "quorum sensing", "surface-binding", "toxic")) %>% 
  setNames(c("Name", "DOI", "Type", "Reproducibility standard", "Web server", "Web server activity", "Model repository", 
             "Training repository", "Antiangiogenic", "Antibacterial", "Antibiofilm", "Anticancer", "Antifungal", "Antihypertensive", 
             "Antiinflammatory", "Antimicrobial", "Antiparasitic", "Antiviral", "Blood-brain barrier", "Cell-penetrating", "Chemotactic", 
             "Neuropeptides", "Quorum sensing", "Surface-binding", "Toxic")) 

```

## {.tabset}

### Data

```{r}
datatable(dat, extensions = c('Buttons', 'FixedColumns'), 
          filter = "top", style = "bootstrap",
          rownames = FALSE, escape = FALSE,
          options = list(dom = "Brtip", pageLength = 25,
                         buttons = c("copy", "csv", "excel", "print"),
                         fixedColumns = list(leftColumns = 1)))
```

### Manual curation procedure

We manually curated all information for software featured in a 'Data' tab. To do that we carefully analysed publications/pre-prints describing given software looking for the following information: 

- peptide activities predicted by a model, 
- link to the web server (**Web server** column), 
- link to the model, i.e. a repository with a trained model which can be used for prediction or an address of the model implemented as standalone software (**Model repository** column),
- link to the repository with all the code and data necessary to train/retrain the model (**Training repository** column).

Then, we checked if the links to the web servers provided in the articles were still working. This information is indicated in the **Web server activity** column. We also inspected available code to determine the reproducibility standard (as described in [Heil et al.](https://doi.org/10.1038/s41592-021-01256-7)) and indicated in the **Reproducibility standard** column). We are additionally using the category *below bronze*, when a model does not fulfill criteria  even for the bronze category.



