---
title: The dynamic landscape of peptide function prediction
output:                                                                                                         
  html_document:
    theme: paper                                                                 
---      

<style type="text/css">
.main-container {
max-width: 1800px;
margin-left: auto;
margin-right: auto;
}
</style>


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(DT)
library(dplyr)

dat <- readRDS("../data/curated_dat.RDS") %>% 
  filter(isPEP == TRUE & isML == TRUE) %>% 
  mutate(DOI = case_when(!is.na(publication) ~ publication,
                         !is.na(preprint) ~ preprint,
                         !is.na(thesis) ~ paste0("<a href='", thesis,"' >Thesis</a>")),
         Type = case_when(!is.na(publication) ~ "Publication",
                          !is.na(preprint) ~ "Pre-print",
                          !is.na(thesis) ~ "Thesis"),
         webserver = sapply(unlist(.[["webserver"]]), function(i) ifelse(i != FALSE, paste0("<a href='", i,"' >Web server</a>"), paste0("false"))),
         modelrepository = sapply(unlist(.[["modelrepository"]]), function(i) ifelse(i != FALSE, paste0("<a href='", i,"' >Repository</a>"), paste0("false"))),
         trainingrepository = sapply(unlist(.[["trainingrepository"]]), function(i) ifelse(i != FALSE, paste0("<a href='", i,"' >Repository</a>"), paste0("false"))))   %>% 
  select(c("name", "DOI", "Type", "reproducibility", "webserver", "webserveractive", "webserverfunctional", "modelrepository", "trainingrepository",
           "antiangiogenic", "antibacterial", "antibiofilm", "anticancer", "antifungal", "antihypertensive", "antiinflammatory",
           "antimicrobial", "antiparasitic", "antiviral", "blood-brain barrier", "cell-penetrating", "chemotactic", "neuropeptides",
           "quorum sensing", "surface-binding", "toxic")) %>% 
  setNames(c("Name", "DOI", "Type", "Reproducibility standard", "Web server", "Web server activity", "Web server functionality", "Model repository", 
             "Training repository", "Antiangiogenic", "Antibacterial", "Antibiofilm", "Anticancer", "Antifungal", "Antihypertensive", 
             "Antiinflammatory", "Antimicrobial", "Antiparasitic", "Antiviral", "Blood-brain barrier", "Cell-penetrating", "Chemotactic", 
             "Neuropeptides", "Quorum sensing", "Surface-binding", "Toxic")) 

```

## {.tabset}

### Software information

```{r}
datatable(dat, extensions = c('Buttons', 'FixedColumns'), 
          filter = "top", style = "bootstrap",
          rownames = FALSE, escape = FALSE,
          options = list(dom = "Brtilp", pageLength = 25,
                         buttons = c("copy", "csv", "excel", "print"),
                         autoWidth = TRUE,
                         # scrollX=TRUE,
                         # columnDefs = list(list(width = '200px', targets = c(0))),
                         fixedColumns = list(leftColumns = 1),
                         filter = list(clear = FALSE, plain = TRUE),
                         lengthMenu = list(25, 50, 75, 100, 125, 150))) %>%  
  formatStyle(c("Web server activity", "Web server functionality", "Antiangiogenic", "Antibacterial", "Antibiofilm", "Anticancer",
                "Antifungal", "Antihypertensive", "Antiinflammatory", "Antimicrobial", "Antiparasitic",
                "Antiviral", "Blood-brain barrier", "Cell-penetrating", "Chemotactic",
                "Neuropeptides", "Quorum sensing", "Surface-binding", "Toxic"),
              backgroundColor = styleEqual(c(FALSE, TRUE), c('#ededed', '#d6edff')))  %>%
  formatStyle(c("Web server", "Model repository", "Training repository"),
              backgroundColor = styleEqual("false", '#ededed')) 
```

### Data collection & curation

To collect software for the prediction of peptide activities, we searched PubMed, Google Scholar and BioRxiv databases using queries such as "antimicrobial peptide prediction", "anticancer peptide prediction", etc. We focused on the search results that described software or reviewed/compared available tools for the prediction of a given activity. We included articles published until 1st July 2022. 

We manually curated all information for software featured in a 'Software information' tab. To do that we carefully analysed publications/pre-prints describing given software looking for the following information: 

- peptide activities predicted by a model, 
- link to the web server (**Web server** column), 
- link to the model, i.e. a repository with a trained model which can be used for prediction or an address of the model implemented as standalone software (**Model repository** column),
- link to the repository with all the code and data necessary to train/retrain the model (**Training repository** column).

Then, we checked if the links to the web servers provided in the articles were still working and if they function correctly, i.e. provide understandable output after running prediction. This information is indicated in the **Web server activity** and **Web server functionality** columns, respectively. We also inspected available code to determine the reproducibility standard (as described in [Heil et al.](https://doi.org/10.1038/s41592-021-01256-7) and indicated in the **Reproducibility standard** column). We are additionally using the category *below bronze*, when a model does not fulfill criteria  even for the bronze category.



