---
title: The dynamic landscape of peptide function prediction
output:                                                                                                         
  html_document:
    theme: paper                                                                 
---      

<style type="text/css">
.main-container {
max-width: 1800px;
margin-left: auto;
margin-right: auto;
}
</style>


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(DT)
library(dplyr)

dat <- readRDS("../data/curated_dat.RDS") %>% 
  filter(isPEP == TRUE & isML == TRUE & !is.na(publication)) %>% 
  mutate(DOI = publication,
         webserver = sapply(unlist(.[["webserver"]]), function(i) ifelse(i != FALSE, paste0("<a href='", i,"' >Web server</a>"), paste0("false"))),
         modelrepository = sapply(unlist(.[["modelrepository"]]), function(i) ifelse(i != FALSE, paste0("<a href='", i,"' >Repository</a>"), paste0("false"))),
         trainingrepository = sapply(unlist(.[["trainingrepository"]]), function(i) ifelse(i != FALSE, paste0("<a href='", i,"' >Repository</a>"), paste0("false"))))   %>% 
  select(c("name", "DOI", "reproducibility", "webserver", "webserveractive", "webserverfunctional", "modelrepository", "trainingrepository",
           "antiangiogenic", "antibacterial", "antibiofilm", "anticancer", "antifungal", "antihypertensive", "antiinflammatory",
           "antimicrobial", "antiparasitic", "antiviral", "blood-brain barrier", "cell-penetrating", "chemotactic", "neuropeptides",
           "quorum sensing", "surface-binding", "toxic")) %>% 
  setNames(c("Name", "DOI", "Reproducibility standard", "Web server", "Web server activity", "Web server functionality", "Model repository", 
             "Training repository", "Antiangiogenic", "Antibacterial", "Antibiofilm", "Anticancer", "Antifungal", "Antihypertensive", 
             "Antiinflammatory", "Antimicrobial", "Antiparasitic", "Antiviral", "Blood-brain barrier", "Cell-penetrating", "Chemotactic", 
             "Neuropeptides", "Quorum sensing", "Surface-binding", "Toxic")) 

hover_text <- c("Name of a software or author if software do not have name",
                "DOI for publications",
                "Reproducibility standard according to Heil et al. 2021. We are additionally using the category below bronze, when a model does not fulfill criteria even for the bronze category.",
                "Link to the web server",
                "Indication if a web server is active, i.e. link provided in the publication is working",
                "Indication if a web server is functional, i.e. provides understandable output after running prediction",
                "Link to the model, i.e. a repository with a trained model which can be used for prediction or an address of the model implemented as standalone software",
                "Link to the repository with all the code and data necessary to train/retrain the model",
                "Indication if the model predicts antiangiogenic activity of peptides",
                "Indication if the model predicts antibacterial activity of peptides",
                "Indication if the model predicts antibiofilm activity of peptides",
                "Indication if the model predicts anticancer activity of peptides",
                "Indication if the model predicts antifungal activity of peptides",
                "Indication if the model predicts antihypertensive activity of peptides",
                "Indication if the model predicts antiinflammatory activity of peptides",
                "Indication if the model predicts antimicrobial activity of peptides",
                "Indication if the model predicts antiparasitic activity of peptides",
                "Indication if the model predicts antiviral activity of peptides",
                "Indication if the model predicts peptides crossing blood-brain barrier",
                "Indication if the model predicts cell-penetrating activity of peptides",
                "Indication if the model predicts chemotactic activity of peptides",
                "Indication if the model predicts neuropeptides",
                "Indication if the model predicts quorum sensing activity of peptides",
                "Indication if the model predicts surface-binding activity of peptides",
                "Indication if the model predicts toxic activity of peptides")

headerCallback <- c(
  "function(thead, data, start, end, display){",
  "  var ncols = data[0].length;",
  sprintf("  var shortnames = [%s]", 
          paste0(paste0("'",colnames(dat),"'"), collapse = ",")),
  sprintf("  var tooltips = [%s];", 
          paste0(paste0("'", hover_text,"'"), collapse = ",")),
  "  for(var i=0; i<ncols; i++){",
  "    $('th:eq('+i+')',thead).attr('title', tooltips[i]).text(shortnames[i]);",
  "  }",
  "}"
)

```

## {.tabset}

### Software information

```{r}
datatable(dat, extensions = c('Buttons', 'FixedColumns'), 
          filter = "top", style = "bootstrap",
          rownames = FALSE, escape = FALSE,
          options = list(dom = "Brtilp", pageLength = 25,
                         buttons = c("copy", "csv", "excel", "print"),
                         autoWidth = TRUE,
                         # scrollX=TRUE,
                         # columnDefs = list(list(width = '200px', targets = c(0))),
                         fixedColumns = list(leftColumns = 1),
                         filter = list(clear = FALSE, plain = TRUE),
                         lengthMenu = list(25, 50, 75, 100, 125, 150),
                         headerCallback = JS(headerCallback))) %>%  
  formatStyle(c("Web server activity", "Web server functionality", "Antiangiogenic", "Antibacterial", "Antibiofilm", "Anticancer",
                "Antifungal", "Antihypertensive", "Antiinflammatory", "Antimicrobial", "Antiparasitic",
                "Antiviral", "Blood-brain barrier", "Cell-penetrating", "Chemotactic",
                "Neuropeptides", "Quorum sensing", "Surface-binding", "Toxic"),
              backgroundColor = styleEqual(c(FALSE, TRUE), c('#ededed', '#d6edff')))  %>%
  formatStyle(c("Web server", "Model repository", "Training repository"),
              backgroundColor = styleEqual("false", '#ededed', default = '#d6edff')) 
```

### Data collection & curation

To collect software for the prediction of peptide activities, we searched PubMed and Google Scholar databases using queries such as "antimicrobial peptide prediction", "anticancer peptide prediction", etc. We focused on the search results that described software or reviewed/compared available tools for the prediction of a given activity. 

Eligibility criteria:

 - the tool needs to be publishen in a peer-review journal until 1st July 2022.
 - we do not include models that are superceded by their next version. For example, we include AmpScanner 2.0, but not AmpScanner.

We manually curated all information for software featured in a 'Software information' tab. To do that we carefully analysed publications/pre-prints describing given software looking for the following information: 

- peptide activities predicted by a model, 
- link to the web server (**Web server** column), 
- link to the model, i.e. a repository with a trained model which can be used for prediction or an address of the model implemented as standalone software (**Model repository** column),
- link to the repository with all the code and data necessary to train/retrain the model (**Training repository** column).

The availablity of web servers was assessed on XYZ.

Then, we checked if the links to the web servers provided in the articles were still working and if they function correctly, i.e. provide understandable output after running prediction. This information is indicated in the **Web server activity** and **Web server functionality** columns, respectively. We also inspected available code to determine the reproducibility standard (as described in [Heil et al.](https://doi.org/10.1038/s41592-021-01256-7) and indicated in the **Reproducibility standard** column). We are additionally using the category *below bronze*, when a model does not fulfill criteria  even for the bronze category.

### About and citation

This website accompanies article: *The dynamic landscape of peptide function prediction* by (in alphabetical order) Oriol Bárcenas, Michał Burdukiewicz, Carlos Pintado-Grima, Katarzyna Sidorczuk and Salvador Ventura.

**Citation**: coming soon

### Authors

#### Oriol Bárcenas

#### Michał Burdukiewicz

Michał Burdukiewicz received Ph.D. in biochemistry from the University of Wrocław in 2019. He is currently working as a post-doc at the Institute of Biotechnology and Biomedicine at the Autonomous University of Barcelona and a research assistant in the Centre for Clinical Research at the Medical University of Białystok. His research interests cover machine learning applications in the functional analysis of peptides and proteins, focusing on amyloids. Moreover, he is co-developing tools for proteomics, mainly hydrogen-deuterium exchange monitored by mass spectrometry.

Contact: michalburdukiewicz[at]gmail.com

Twitter: https://twitter.com/burdukiewicz

Website: https://github.com/michbur

#### Carlos Pintado-Grima

#### Katarzyna Sidorczuk

Katarzyna Sidorczuk received the M.Sc. degree in biotechnology from the University of Wrocław, Poland, in 2019. She is currently pursuing the Ph.D. degree in biological sciences at the University of Wrocław. Her research focuses on bioinformatics and machine learning approaches for the analysis and prediction of peptide functions, protein targeting sequences and bacterial adhesins.

#### Salvador Ventura
